---
title: "collecting_enviromental_data"
author: "Katie Wang"
date: "2024-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, messages = FALSE)
```

```{r}
#loading in the packages :D
library(tidyverse)
library(sdmpredictors)
library(raster)
library(sp)
library(terra)
library(sf)
library(stars)
library(here)
library(dplyr)
```

```{r}
#listing all of the different variables 
library(biooracler)

dataset <- list_layers()
dataset
```

```{r}
latitude = c(-89.975, 89.975)
longitude = c(-179.975, 179.975)
constraints = list(latitude, longitude)
names(constraints) = c("latitude", "longitude")
```

```{r}
info_layer("thetao_ssp119_2020_2100_depthmean")
```

```{r}
#importing the mean ocean temperature
dataset_id1 <- "thetao_ssp119_2020_2100_depthmean"
variables = c("thetao_mean") #choosing the mean of the mean temperature 
mean_ocean_temp <- download_layers(dataset_id1, variables, constraints)
```

```{r}
#importing the mean salinity
dataset_id3<- "so_baseline_2000_2019_depthmean"
variables = c("so_mean") #choosing the mean of salinity
salinity <- download_layers(dataset_id3, variables, constraints)
```

```{r}
#importing the mean TotalPhytoplankton
dataset_id4 <- "phyc_baseline_2000_2020_depthmean"
variables = c("phyc_mean") #choosing the mean of TotalPhytoplankton <- can be rep of primary productivity
totalphytoplankton <- download_layers(dataset_id4, variables, constraints)
```

```{r}
#importing the mean pH
dataset_id5 <- "ph_baseline_2000_2018_depthmean"
variables = c("ph_mean") #choosing the mean of pH
ph <- download_layers(dataset_id5, variables, constraints)
```

```{r}
#importing the mean chlorophyll
dataset_id6 <- "chl_baseline_2000_2018_depthmean"
variables = c("chl_mean") #choosing the mean of chlorophyll
chlorophyll <- download_layers(dataset_id6, variables, constraints)
```

```{r}
#importing the MixedLayerDepth
dataset_id7 <- "mlotst_baseline_2000_2019_depthsurf"
variables = c("mlotst_mean") #choosing the mean of MixedLayerDepth
mixedlayerdepth <- download_layers(dataset_id7, variables, constraints)
```

Can plot(variable) <- to see the current data

```{r}
#reading the manta data 
manta_data <- st_read("Data/manta_csr.shp")
```


```{r}
#setting the max ocean temp crs to be the same
mean_temp <- project(mean_ocean_temp,
                                 crs(manta_data))
terra::crop(mean_ocean_temp, manta_data)
temp = terra::extract(x = mean_ocean_temp,y = manta_data)

temp <- temp %>%
  filter(!is.na(thetao_mean_1) & !is.na(thetao_mean_2)) %>% 
  dplyr::select("thetao_mean_2") %>% 
  rename("thetao_mean2" = "mean_ocean_temp")
```

```{r}
mean_salinity <- project(salinity,
                                 crs(manta_data))
terra::crop(salinity, manta_data)
salinity = terra::extract(x = salinity,y = manta_data)

salinity <- salinity %>%
  filter(!is.na(so_mean_1) & !is.na(so_mean_2)) %>% 
  dplyr::select("so_mean_2") %>% 
  rename("so_mean_2" = "mean_salinity")
```


```{r}
#setting the mean total phytoplankton crs to be the same
mean_totalphytoplankton <- project(totalphytoplankton,
                                 crs(manta_data))
terra::crop(totalphytoplankton, manta_data)
totalphytoplankton = terra::extract(x = totalphytoplankton,y = manta_data)

totalphytoplankton <- totalphytoplankton %>%
  filter(!is.na(phyc_mean_1) & !is.na(phyc_mean_2)) %>% 
  dplyr::select("phyc_mean_2") %>% 
  rename("phyc_mean_2" = "mean_total_phytoplankton")
```


```{r}
#setting the mean ph crs to be the same
mean_ph <- project(ph,
                                 crs(manta_data))
terra::crop(ph, manta_data)
ph = terra::extract(x = ph,y = manta_data)
ph <- ph %>%
  filter(!is.na(ph_mean_1) & !is.na(ph_mean_2)) %>% 
  dplyr::select("ph_mean_2") %>% 
  rename("ph_mean_2" = "mean_ph")

```

```{r}
#setting the mean chlorophyll crs to be the same
mean_chlorophyll <- project(chlorophyll,
                                 crs(manta_data))
terra::crop(chlorophyll, manta_data)
chlorophyll = terra::extract(x = chlorophyll,y = manta_data)
chlorophyll <- chlorophyll %>%
  filter(!is.na(cl_mean_1) & !is.na(cl_mean_2)) %>% 
  dplyr::select("thetao_mean_2") %>% 
  rename("thetao_mean_2" = "mean_chlorophyll")
```


```{r}
#setting the Mixed Layer Depth crs to be the same
mean_MixedLayerDepth <- project(MixedLayerDepth,
                                 crs(manta_data))
terra::crop(MixedLayerDepth, manta_data)
MixedLayerDepth = terra::extract(x = MixedLayerDepth,y = manta_data)
MixedLayerDepth <- MixedLayerDepth %>%
  filter(!is.na(mlotst_mean_1) & !is.na(mlotst_mean_2)) %>% 
  dplyr::select("nlotst_mean_2") %>% 
  rename("nlotst_mean_2" = "mean_MixedLayerDepth")
```

```{r}
write.csv("Data/temp")
```

```{r}
write.csv("Data/salinity")
```

```{r}
write.csv("Data/totalphytoplankton")
```

```{r}
write.csv("Data/ph")
```

```{r}
write.csv("Data/chlorophyll")
```

```{r}
write.csv("Data/mean_MixedLayerDepth")
```

```{r}
temp_sf <- st_as_sf(temp, coords = c("longitude", "latitude"), crs = 4326)

p1 <- ggplot(temp_sf) +
  geom_sf(aes(fill = thetao_mean_1)) +
  scale_fill_viridis_c(name = "Mean Temperature (°C)") +
  labs(x = "Longitude", y = "Latitude", title = "Mean Temperature in Each MPA",
       fill = "Mean Temperature") +
  theme_minimal()

# Display the plot
print(p1)
```

```{r}
glm(formula = present ~ tmin + precip, family = binomial(link = "logit"), 
     data = hooded_warb_data)
```

